---
title: "p8105_hw2_qy2234"
author: "Michael Yan"
date: "9/25/2019"
output: github_document
---

## Problem 1

```{r Question_one_part_one}
## load the tidyverse and readxl package
library(tidyverse)
library(readxl)

## specify the sheet in the Excel file and to omit non-data entries
tw_data = 
  read_excel(
  "./P1_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1) %>%
  janitor::clean_names() %>% 
  select(-x15) %>% 
  drop_na(dumpster) %>%
  select(dumpster, month, year, everything()) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

```{r Question_one_2017_2018}
## read and clean precipitation data for 2017 and 2018

## clear 2017 data
tw_data_17 = read_excel(
  "./P1_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
  sheet = 4, skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2017") %>% 
  rename(percipitation_in_in = total)

## clear 2018 data
tw_data_18 = read_excel(
  "./P1_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
  sheet = 3, skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2018") %>% 
  rename(percipitation_in_in = total)

## combine 2017 2018 data
tw_data_17_18 = 
  left_join(tw_data_17, tw_data_18, by = "month") %>% 
  drop_na(month) %>%
  mutate(month = as.character(month.name))
```
#### Question_one_description
* The number of observations in tw_data is 
`r nrow(tw_data)`
* The number of observations in tw_data_17_18 is 
`r nrow(tw_data_17_18)`
* Some key variables in tw-data include 
`r colnames(tw_data)`
* Some key variables in tw_data_17_18 include 
`r colnames(tw_data_17_18)`
* The total precipitation in 2018 is 
`r sum(pull(tw_data_17_18, percipitation_in_in.y), na.rm = TRUE)`
* The median number of sports balls in a dumpster in 2017 is
`r median(pull(tw_data, dumpster), na.rm = TRUE)`



## Problem 2














