---
title: "p8105_hw2_qy2234"
author: "Michael Yan"
date: "9/25/2019"
output: github_document
---

## Problem 1

```{r Question_one_part_one}
## load the tidyverse and readxl package
library(tidyverse)
library(readxl)

## import data and specify the sheet in the Excel file and to omit non-data entries
tw_data = 
  read_excel(
  "./P1_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1) %>%
  janitor::clean_names() %>% 
  select(-x15,-x16, -x17) %>% 
  drop_na(dumpster) %>%
  select(dumpster, month, year, everything()) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

```{r Question_one_2017_2018}
## read and clean precipitation data for 2017 and 2018

## import and clear 2017 data
tw_data_17 = read_excel(
  "./P1_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 6, skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2017") %>% 
  rename(percipitation_in_in = total)

## import and clear 2018 data
tw_data_18 = read_excel(
  "./P1_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = 5, skip = 1) %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2018") %>% 
  rename(percipitation_in_in = total)

## combine 2017 2018 data
tw_data_17_18 = 
  left_join(tw_data_17, tw_data_18, by = "month") %>% 
  drop_na(month) %>%
  mutate(month = month.name)
```
#### Question_one_description
* The number of observations in tw_data is 
`r nrow(tw_data)`
* The number of observations in tw_data_17_18 is 
`r nrow(tw_data_17_18)`
* Some key variables in tw-data include 
dumpster, month, and year
* Some key variables in tw_data_17_18 include 
month, year
* The total precipitation in 2018 is 
`r sum(pull(tw_data_17_18, percipitation_in_in.y), na.rm = TRUE)`
* The median number of sports balls in a dumpster in 2017 is
`r median(pull(tw_data, dumpster), na.rm = TRUE)`


## Problem 2

```{r Problem_2_part_1}
## load the tidyverse and readxl package
library(tidyverse)
library(readxl)

## import and clean the data in pols-month.csv
pols_month_data = read_csv(
  "./P2_data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), convert = TRUE) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(month = str_to_lower(month)) %>% 
  pivot_longer(gov_gop:rep_gop, 
               names_to = "gop_groups",
               values_to = "count_gop") %>% 
  pivot_longer(gov_dem:rep_dem, 
               names_to = "dem_groups",
               values_to = "count_dem") %>% 
  pivot_longer(prez_gop:prez_dem, 
               names_to = "president",
               values_to = "prez_gop_or_dem") %>% 
  select(-president, -day) %>% 
  rename(president = prez_gop_or_dem)
```

```{r Problem_2_part_2}
## import and clean the data in snp.csv
snp_data = read_csv(
  "./P2_data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c("month", "day", "year"), convert = TRUE) %>% 
  mutate(month = month.abb[month]) %>% 
  mutate(month = str_to_lower(month)) %>%
  arrange(year, month) %>% 
  select(year, month, day, close)
```

```{r Problem_2_part_3}
## import and clean the data in unemployment.csv
unemployment_data = read_csv(
  "./P2_data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment") 
```

```{r Problem_2_part_4}
## join the datasets by merging snp into pols, and merging unemployment into the result
pols_snp = 
  left_join(pols_month_data, snp_data, by = "year")
pols_snp_unemployment = 
  left_join(pols_snp, unemployment_data, by = "year")
```
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

#### Question_two_description
* The pols_month_data contains the number of governors, senators, and representatives 
from both parties at a given time. The snp_data provides the closing values of the 
S&P stock index on the associated date. The unemployment_data shows 
the percentage of unemployment in each month of the associated year. Finally, 
the pols_snp_unemployment dataset gives us a view of 
all the information associate with a given time point, 
including the number of governors, senators, and representatives 
from both parties, the S&P stock index, and the percentage of unemployment. 
It also gives us the information about when the measurements were done
(which month, or day, if applicabel).

The dimension of the final resulting dataset is 
`r nrow(pols_snp_unemployment)` by `r ncol(pols_snp_unemployment)`

The range of years is 
`r min(pull(pols_month_data, year))` - `r max(pull(pols_month_data, year))` 

The names of key variables include year and month


## Problem 3

```{r Question_three_part_one}
## import and clean the data in unemployment.csv
baby_name_data = read_csv(
  "./P3_data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  distinct() %>% 
  mutate(gender = str_to_lower(gender)) %>% 
  mutate(ethnicity = str_to_lower(ethnicity)) %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name))
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year. Produce a similar table showing the most popular name among male children over time.

```{r Question_three_part_two}
## produce a well-structured, reader-friendly table for girl's name
baby_name_data_tidy_girl = 
  pivot_wider(
  baby_name_data, 
  names_from = "year_of_birth", 
  values_from = "count")

## produce a well-structured, reader-friendly table for boy's name
baby_name_data_tidy_boy = 
  pivot_wider(
  baby_name_data, 
  names_from = "year_of_birth", 
  values_from = "count") %>% 
  arrange(desc(gender))
```

```{r Question_three_part_three}
## for male, white non-hispanic children born in 2016, produce a scatter plot
data_for_scatter_plot = 
  filter(baby_name_data, year_of_birth == 2016, gender == "male", 
         ethnicity == "white non hispanic")
ggplot(baby_name_data, aes(x = rank, y = count)) + 
  geom_point()
```








